Revealing transient strain in geodetic data with Gaussian process regression

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex

\textbf{Review of the paper : Revealing transient strain in geodetic data
with Gaussian process regression}

\textbf{Overview of the work: The paper focuses on a new non-parametric
Bayesian method to estimate transient strain rate from GNSS data. It
is based on Gaussian Process regression. The method also includes an
outlier removal processing step. This technique is then applied to the
detection of slow slip events in Cascadia. The validity of the
detected transient signals is accomplished via comparison with the
seismic tremor recorded along the Cascadia range.}

\textbf{The article is well-written and your algorithm is very interesting. I
think the use of Gaussian Process Regression is novel and open new
roads in GNSS data analysis. I believe that you spend a lot of efforts
in trying to justify most of the assumptions behind your model,
together with the effects of using different kind of prior. Perhaps,
that overshadows slightly your results. You could consider to
reorganize the article to make it more readable for geophysicists.}

I have taken several steps to make my revised draft more accessible to
someone who does not have a strong background in stochastic processes:
I rewritten the methods section to be clearer, contain more examples,
and elaborate on potentially unfamiliar concepts. I have also added
more references to the methods section to make it easier for the
reader to find additional information. I sincerely hope that my
revised draft is more readable to geophysicists.

@@

\textbf{About the method:}

\textbf{The method is well established. The assumptions to justify the various
hypothesis on the deterministic models of the GNSS time series are
generally well supported (seasonal signal, tectonic rate, offsets).}

\textbf{1- Now, the correlation of the stochastic processes is not completely
clear. When the authors justify the model behind equation 3, wij is a
normally distributed, uncorrelated noise. In statistic, it is equal to
a WGN process. Then, the authors introduce a separate parameter (eta)
which models the temporally correlated noise. This parameter is also
following a Gaussian distribution, with a zero-mean and a covariance
matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
modelling GNSS time series. The standard is based on Williams 2003,
where the author established a stochastic model with a covariance
matrix as sum of identity matrix for the white noise and another
matrix which represents the coloured noise. The coloured noise
covariance matrix is defined differently (see the literature – Bock and
Melgar Reviews of Geophysics 2016, He et al., Journal of geodynamics
2017).}

I have changed my description of the data (eq. 3 in the revised draft)
so that it includes a single term, $\eta$, that encompasses correlated
and uncorrelated noise. I then break $\eta$ down into correlated and
uncorrelated components in Section 3.1 (eq. 25). The covariance matrix
for $\eta$ in eq. 26 is more consistent with the literature.

@@

\textbf{2- First, the author should give a summary of stochastic noise
modelling in GNSS. I also think that a discussion is needed to relate
to previous models. Why using separated stochastic models? Also I
would comment on the possibility that the (low) spatial correlation
between the parameters in your deterministic model (i.e. secular
velocities) could be absorbed in the estimation of the covariance
matrix $C_{\eta}$.}

I believe the reviewer is refering to page 4-5 of the original draft,
where I first introduce $\eta$. I do not think this is an appropriate
place for a discussion on the noise model, or for a summary of past
studies. The paper is laid out so that Section 2 contains a general
description of the method for estimating transient strain rates, and
Section 3 contains details that are specific to the particular
application. There is no universally appropriate noise model so I
decided to save a discussion on the noise model for Section 3. This is
clarified in Section 2 of the revised draft when I say "The
appropriate noise model may vary depending on the application, and we
hold off on specifying the covariance matrix, $\mitbf{C}_{\eta_i}$,
until Section \ref{sec:NoiseModel}."

@@

\textbf{Now, I am confused When looking at p.8 Section 4.1, the
description of the noise models should be given as a subsection of
Section 2 p.4. I would also improve the literature review. Several
other models have also been discussed (i.e. Montillet et al., 2014
uses a fractional Brownian motion model; ARMA, ARFIMA, GGM or Band
pass noise – see Het et al., 2017).}

Again, we think that the description of the noise model should remain
in Section 3 because it is specific to the Pacific Northwest
application.

We have added a few more references in Section 3.1 of the revised
draft TODO to emphasize that GNSS noise modeling is a well-worked
area. We refer the reader to Bock and Melgar 2016 and He et al 2017
for a more thorough review.

@@

\textbf{Outlier detection:}

\textbf{1- P7 : I would give a longer summary in the introduction about all
the efforts in outlier detection in GNSS time series. The first
paragraph p.7 is not enough.}

In Section 2.2 of the revised draft, I have expanded the discussion of
previous efforts to detect outliers in GNSS data.

@@

\textbf{p.2 line 43 :”too large of an area” ... perhaps “ a very large area”}

Done

\textbf{P5 line 32 : “formal data uncertainty”: can you define it?}

By "formal data uncertainties" I meant "the derived uncertainties that
accompany the GNSS displacement solutions". The revised text has been
clarified.

@@

\textbf{P 7 line 54-60: This paragraph can be further enhanced with references
(Bock and Melgar, 2016, Gazeaux et al., 2013).}

I have added a reference to Gazeaux et al. 2013.

\textbf{P 8 : first paragraph. add references on Cascadia range (Aguiar,
Melbourne, Scriver 2009), (Melbourne and Szeliga 2005), (Melbourne and
Webb, 2003, 2002).}

We agree that we need to acknowledge the extensive amount of research
on Cascadia SSEs. In the revised draft, we refer the reader to the
review by Schartz and Rokosky 2007.

@@

\textbf{P.8 line 20 delete www.unavco.org (only in Acknowledgment)}

Done

\textbf{p. 11 line 45 “common mode error” you need to add some references and
define it properly.}

In Section 3.1 of the revised draft, where we elaborate on the noise
model, we describe common mode error by saying "Another significant
source of noise in GNSS data is common mode error (e.g., Wdowinski et
al. 1997; Dong et al. 2006), which is noise that is highly spatially
correlated." We believe this definition and these references are
sufficient.

@@

\textbf{p.13 : “Wendland functions have compact support and hence their
corresponding covariance.. sparse”. I would just mention that the
associated covariance matrices are sparse.}

I recognize that it may be a bit redundant to say that the covariance
function is compact and thus the covariance matrices are sparse.
However, Wendland covariance functions are typically described as
"compact" in the literature, and I think it is a key adjective that
should not be dropped out. I have left the wording as is.

@@

\textbf{p.13 line 22 what is an “isotropic Gaussian process”? please add a
definition.}

An Gaussian process is isotropic if its covariance function,
$C(x,x')$, can be written as a function of $||x - x'||_2^2$. An
isotropic Gaussian process is also stationary, meaning that its
statistical properties are invariant to translations. We have replace
"isotropic" with "stationary" in the text because "stationary" is a
more appropriate, and perhaps more familiar, term.

@@

\textbf{P.20 : “ Using a compact covariance function ..” Again I would just
say that the covariance matrix is sparse and that s why you have the
mathematical simplification.}

My personal preference is to leave the wording as is.

\textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
60 “prohibitive when using sveral years of daily GNSS”. It would
interesting to have some number. How do you quantify the computation
time to process longer and longer time series?}

I agree that numerical values would be interesting; however, I am
reluctant to quantify the computational cost of evaluation transient
strain rates. There are many factors that influence the computational
cost (whether sparse matrices are used, the type of sparse matrices
that are used, the linear solver algorithm, etc.) and I am not
convinced that my implementation is the most efficient. So I do not
give details on the computation time, but I note in the discussion
that this is an area for further research.

### 2017-11-10 ResponseToReviewers/ResponseToReviewers.tex

\textbf{1- Now, the correlation of the stochastic processes is not completely
clear. When the authors justify the model behind equation 3, wij is a
normally distributed, uncorrelated noise. In statistic, it is equal to
a WGN process. Then, the authors introduce a separate parameter (eta)
which models the temporally correlated noise. This parameter is also
following a Gaussian distribution, with a zero-mean and a covariance
matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
modelling GNSS time series. The standard is based on Williams 2003,
where the author established a stochastic model with a covariance
matrix as sum of identity matrix for the white noise and another
matrix which represents the coloured noise. The coloured noise
covariance matrix is defined differently (see the literature – Bock and
Melgar Reviews of Geophysics 2016, He et al., Journal of geodynamics
2017).}

I have changed my description of the data (eq. 3 in the revised draft)
so that it includes a single term, $\eta$, that encompasses correlated
and uncorrelated noise. I then break $\eta$ down into correlated and
uncorrelated components in Section 3.1 (eq. 25). The covariance matrix
for $\eta$ in eq. 26 is more consistent with the literature.

\textbf{2- First, the author should give a summary of stochastic noise
modelling in GNSS. I also think that a discussion is needed to relate
to previous models. Why using separated stochastic models? Also I
would comment on the possibility that the (low) spatial correlation
between the parameters in your deterministic model (i.e. secular
velocities) could be absorbed in the estimation of the covariance
matrix $C_{\eta}$.}

I believe the reviewer is refering to page 4-5 of the original draft,
where I first introduce $\eta$. I do not think this is an appropriate
place for a discussion on the noise model, or for a summary of past
studies. The paper is laid out so that Section 2 contains a general
description of the method for estimating transient strain rates, and
Section 3 contains details that are specific to the particular
application. There is no universally appropriate noise model so I
decided to save a discussion on the noise model for Section 3. This is
clarified in Section 2 of the revised draft when I say "The
appropriate noise model may vary depending on the application, and we
hold off on specifying the covariance matrix, $C_{\eta_i}$,
until Section 3.1."

### 2017-11-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

In this section we estimate transient strain rates in the Pacific
Northwest, and we are specifically interested in identifying transient
strain resulting from SSEs \citep[e.g.,][]{Dragert2001}. Before
estimating transient strain rates, we establish a noise model for GNSS
stations in this region, and we establish a prior Gaussian process to
describe displacements from SSEs. SSEs in the Pacific Northwest can be
detected by monitoring for associated seismic tremor
\citep{Rogers2003}, which is actively being done by the Pacific
Northwest Seismic Network \citep{Wech2010}. We can then compare the
tremor records to our estimated transient strain rates to verify that
we are indeed identifying strain from SSEs.

@@

The temporally correlated noise in GNSS data has been thoroughly
studied over the past two decades \citep[e.g.,][]{Zhang1997, Mao1999,
Williams2004, Langbein2008}. In these studies, temporally correlated
noise tends to be described with some combination of Brownian motion
(also known as random walk noise or a Weiner process), a first-order
Gauss-Markov (FOGM) process, and/or flicker noise. There is some
physical justification for using Brownian motion as a noise model
because it accurately describes the power spectrum of motion resulting
from instability in a geodetic monument \citep[e.g.,][]{Wyatt1982,
Wyatt1989}. In particular, the power spectrums for Brownian motion and
monument motion both decay proportionally to $f^{-2}$, where $f$ is
frequency. However, Brownian motion necessarily contains a reference
time at which the process begins. Since there is no notion of when
noise ``begins" in GNSS data, we do not find Brownian motion to be an
appropriate noise model. On the other hand, a FOGM process has no
reference time (i.e., it is stationary), and its power spectrum,
\begin{equation}\label{eq:FOGMPS}
S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2},
\end{equation}
decays proportionally to $f^{-2}$ above the cutoff frequency $\alpha /
(2 \pi)$. We then choose $z_i$ to be a spatially
uncorrelated FOGM process, which has the covariance function
\begin{equation}\label{eq:FOGM}
C_{z_i}\left((\pos,t),(\pos',t')\right) = 
\frac{\beta^2}{2\alpha}
\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'},
\end{equation}
where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise.
See \citet[sec. B2]{Rasmussen2006} for more information on the FOGM
process.

@@

We consider three potential choices for $T_i$. The first option is the
one-dimensional SE covariance function,
\begin{equation}\label{eq:TimeSE}
T_i(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
\end{equation}
Note that $T_i$ includes the parameter $\phi$, which serves to scale
the covariance function $C_{u_i}$. The second option is a member of
the Wendland class of covariance functions \citep{Wendland2005}.
Wendland covariance functions have compact support and hence their
corresponding covariance matrices are sparse. In our analysis, we
exploit this sparsity with the CHOLMOD software package
\citep{Chen2008}. Wendland covariance functions can be constructed
such that they are positive definite in $\mathbb{R}^d$ and their
corresponding Gaussian process can be differentiated $k$ times. We use
$d=1$, because we are describing the temporal covariance of $u_i$, and
we use $k=2$, giving samples of $u_i$ continuous velocities and
accelerations. The corresponding Wendland covariance function is
\begin{equation}\label{eq:Wendland} 
T_i(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^5_+ 
            \left(\frac{8|t - t'|^2}{\tau^2} + \frac{5|t - t'|}{\tau} + 1\right), 
\end{equation} 
where $(t)_+ = \mathrm{max}(0,t)$. The third option for $T_i$ is the
covariance function for integrated Brownian motion (IBM). IBM is a
Gaussian process with zero mean and its covariance function can be
found by integrating the covariance function for Brownian motion,
\begin{align}\label{eq:IBM}
T_i(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
          &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
\end{align}
IBM has been used in the context of Kalman filtering as a
non-parametric model for the time dependence of geophysical signals
\citep[e.g.,][]{Segall1997, McGuire2003, Ohtani2010, Hines2016a}.
Similar to Brownian motion, IBM has a reference time, $t=0$, at which
the process begins. For some geophysical signals, it is appropriate to
have this reference time. For example, if we are trying to identify
postseismic deformation then $t$ should be zero at the time of the
earthquake.  However, if we are interested in detecting transient
events, where there is no known start time, then IBM may not be an
appropriate prior model. Instead, one may prefer to use the SE or
Wendland covariance functions because they are stationary. In the
following analysis, we make the choice that $t$ is zero on the first
epoch of $\data_i^*$. Using an earlier reference time does not change
the results discussed in this section.

@@

We use the REML method to determine the appropriate values for the
parameters in $C_{u_i}$, which are $\ell$, $\phi$, and $\tau$. We
refer to these parameters collectively as $\mitbf{\theta}$. Just as in
Section \ref{sec:NoiseModel}, we pick $\mitbf{\theta}$ to maximize the
probability of drawing $\data_i^*$ from $\data_i$, but now we are
including $u_i(\points)$ in $\data_i$. and we are optimizing the
parameters for $C_{u_i}$ rather than $\mitbf{C}_{\eta_i}$. We divide
the GNSS data into seven subsets that are four months long and each
centered on the time of a SSE. The times of the seven SSEs are
determined with tremor records from \cite{Wech2010}. We find the
optimal $\mitbf{\theta}$ for each subset of data, for each
displacement component, and for each choice of $T_i$. The REML
likelihood function that we are maximizing with respect to
$\mitbf{\theta}$ is now
\begin{equation}\label{eq:REMLPrior} 
\mathcal{L}(\mitbf{\theta}) =
\left(\frac{\left| \G^T \G \right|}{(2\pi)^{n-6m}
                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
                                    \left| \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2} \data_i^{*T} \mitbf{K}(\mitbf{\theta}) \data_i^*},
\end{equation}
where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
                            \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G
                            \left( \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right)^{-1}
                            \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1}
\end{equation}
and $\mitbf{\Sigma}(\mitbf{\theta}) = \mitbf{C}_{\eta_i} +
C_{u_i}(\points,\points;\mitbf{\theta})$. In the above equation,
$\data^*_i$ consists of a single subset of data. The estimated
parameters are summarized in Table 1. Based on the interquartile
ranges, the estimated parameters for the SE and Wendland covariance
functions do not vary significantly between SSEs. This suggests that
the median of the estimated parameters should be appropriate for all
the Pacific Northwest SSEs. For the IBM model, there are several
anomalously large values for $\ell$ and $\phi$, which is why the
interquartile ranges are large.

>>>>

In this section we estimate transient strain rates in the Pacific
Northwest, and we are specifically interested in identifying transient
strain resulting from SSEs. SSEs in the Pacific Northwest were first
discovered by \citep[e.g.,][]{Dragert2001} and we refer the reader to
\citet{Schwartz2007} for a review of the subsequent research. Before
estimating transient strain rates, we establish a noise model for GNSS
stations in this region, and we establish a prior Gaussian process to
describe displacements from SSEs. SSEs in the Pacific Northwest can be
detected by monitoring for associated seismic tremor
\citep{Rogers2003}, which is actively being done by the Pacific
Northwest Seismic Network \citep{Wech2010}. We can then compare the
tremor records to our estimated transient strain rates to verify that
we are indeed identifying strain from SSEs.

@@

The temporally correlated noise in GNSS data has been thoroughly
studied over the past two decades \citep[e.g.,][]{Zhang1997, Mao1999,
Williams2004, Langbein2008} (See \citet{Bock2016} or \citet{He2017}
for a review of GNSS noise modelling). In these studies, temporally
correlated noise tends to be described with some combination of
Brownian motion (also known as random walk noise or a Weiner process),
a first-order Gauss-Markov (FOGM) process, and/or flicker noise. There
is some physical justification for using Brownian motion as a noise
model because it accurately describes the power spectrum of motion
resulting from instability in a geodetic monument
\citep[e.g.,][]{Wyatt1982, Wyatt1989}. In particular, the power
spectrums for Brownian motion and monument motion both decay
proportionally to $f^{-2}$, where $f$ is frequency. However, Brownian
motion necessarily contains a reference time at which the process
begins. Since there is no notion of when noise ``begins" in GNSS data,
we do not find Brownian motion to be an appropriate noise model. On
the other hand, a FOGM process has no reference time (i.e., it is
stationary), and its power spectrum, \begin{equation}\label{eq:FOGMPS}
S(f) = \frac{\beta^2}{(2 \pi f)^2 + \alpha^2}, \end{equation} decays
proportionally to $f^{-2}$ above the cutoff frequency $\alpha / (2
\pi)$. We then choose $z_i$ to be a spatially uncorrelated FOGM
process, which has the covariance function
\begin{equation}\label{eq:FOGM}
C_{z_i}\left((\pos,t),(\pos',t')\right) = \frac{\beta^2}{2\alpha}
\exp\left(-\alpha|t - t'|\right)\delta_{\pos,\pos'}, \end{equation}
where $\delta_{\pos,\pos'}$ is 1 if $\pos = \pos'$ and 0 otherwise.
See \citet[sec. B2]{Rasmussen2006} for more information on the FOGM
process.

@@

We consider three potential choices for $T_i$. The first option is the
one-dimensional SE covariance function,
\begin{equation}\label{eq:TimeSE}
T_i(t,t') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right).
\end{equation}

Note that $T_i$ includes the parameter $\phi$, which serves to scale
the covariance function $C_{u_i}$. The second option is a member of
the Wendland class of covariance functions \citep{Wendland2005}.
Wendland covariance functions have compact support and hence their
corresponding covariance matrices are sparse. We exploit this sparsity
with the CHOLMOD software package \citep{Chen2008}. Wendland
covariance functions can be constructed such that they are positive
definite in $\mathbb{R}^d$ and their corresponding Gaussian process
can be differentiated $k$ times. We use $d=1$, because we are
describing the temporal covariance of $u_i$, and we use $k=2$, giving
samples of $u_i$ continuous velocities and accelerations. The
corresponding Wendland covariance function is

\begin{equation}\label{eq:Wendland} 
T_i(t,t') = \phi^2\left(1 - \frac{|t - t'|}{\tau}\right)^5_+ 
            \left(\frac{8|t - t'|^2}{\tau^2} + \frac{5|t - t'|}{\tau} + 1\right), 
\end{equation} 
where $(t)_+ = \mathrm{max}(0,t)$. The third option for $T_i$ is the
covariance function for integrated Brownian motion (IBM). IBM is a
Gaussian process with zero mean and its covariance function can be
found by integrating the covariance function for Brownian motion,
\begin{align}\label{eq:IBM}
T_i(t,t') &= \int_0^t \int_0^{t'} \phi^2 \min(s,s') \,ds'\,ds \\
          &= \frac{\phi^2}{2}\min(t,t')^2 \left(\max(t,t') - \frac{1}{3}\min(t,t')\right), \ \ \ t,t' \geq 0.
\end{align}
IBM has been used in the context of Kalman filtering as a
non-parametric model for the time dependence of geophysical signals
\citep[e.g.,][]{Segall1997, McGuire2003, Ohtani2010, Hines2016a}.
Similar to Brownian motion, IBM has a reference time, $t=0$, at which
the process begins. For some geophysical signals, it is appropriate to
have this reference time. For example, if we are trying to identify
postseismic deformation then $t$ should be zero at the time of the
earthquake.  However, if we are interested in detecting transient
events, where there is no known start time, then IBM may not be an
appropriate prior model. Instead, one may prefer to use the SE or
Wendland covariance functions because they are stationary. In the
following analysis, we make the choice that $t$ is zero on the first
epoch of $\data_i^*$. Using an earlier reference time does not change
the results discussed in this section.

@@

We use the REML method to determine the appropriate values for the
parameters in $C_{u_i}$, which are $\ell$, $\phi$, and $\tau$. We
refer to these parameters collectively as $\mitbf{\theta}$. Just as in
Section \ref{sec:NoiseModel}, we pick $\mitbf{\theta}$ to maximize the
probability of drawing $\data_i^*$ from $\data_i$, but now we are
including $u_i(\points)$ in $\data_i$, and we are optimizing the
parameters for $C_{u_i}$ rather than $\mitbf{C}_{\eta_i}$. We divide
the GNSS data into seven subsets that are four months long and each
centered on the time of a SSE. The times of the seven SSEs are
determined with tremor records from \cite{Wech2010}. We find the
optimal $\mitbf{\theta}$ for each subset of data, for each
displacement component, and for each choice of $T_i$. The REML
likelihood function that we are maximizing with respect to
$\mitbf{\theta}$ is now
\begin{equation}\label{eq:REMLPrior} 
\mathcal{L}(\mitbf{\theta}) =
\left(\frac{\left| \G^T \G \right|}{(2\pi)^{n-6m}
                                    \left| \mitbf{\Sigma}(\mitbf{\theta}) \right| 
                                    \left| \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right|}\right)^{\frac{1}{2}} 
                              e^{-\tfrac{1}{2} \data_i^{*T} \mitbf{K}(\mitbf{\theta}) \data_i^*},
\end{equation}
where
\begin{equation}
\mitbf{K}(\mitbf{\theta}) = \mitbf{\Sigma}(\mitbf{\theta})^{-1} - 
                            \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G
                            \left( \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1} \G \right)^{-1}
                            \G^T \mitbf{\Sigma}(\mitbf{\theta})^{-1}
\end{equation}
and $\mitbf{\Sigma}(\mitbf{\theta}) = \mitbf{C}_{\eta_i} +
C_{u_i}(\points,\points;\mitbf{\theta})$. In the above equation,
$\data^*_i$ consists of a single subset of data. The estimated
parameters are summarized in Table 1. Based on the interquartile
ranges, the estimated parameters for the SE and Wendland covariance
functions do not vary significantly between SSEs. This suggests that
the median of the estimated parameters should be appropriate for all
the Pacific Northwest SSEs. For the IBM model, there are several
anomalously large values for $\ell$ and $\phi$, which is why the
interquartile ranges are large.
